Question No: 1
In an exam, the score records of 16 students (suppose they worked independently) are given below:
45, 88, 58,74, 56, 70, 84, 73, 60, 69, 65, 72, 72, 69, 64, 71 
(a) Plot the data using BoxPlot, are there detected outliers? Indicate their values if any. (5pt)

```{r}
x<-c(45, 88, 58,74, 56, 70, 84, 73, 60, 69, 65, 72, 72, 69, 64, 71)
x
graphics.off()
boxplot(x, ylab = "students", main= "Box plot of the students score")
outlier <- boxplot.stats(x)
outlier

```

(b) Summarize the data (e.g., mean, max, min, mode, median, range, standard deviation, variance)
Then, use appropriate test to decide whether the data are normal. (10pt)

```{r}
summary <- summary(x)
summary
#appropriate method to confirm the test
#mean test
mean_val<- mean(x)
mean_val
#median test
median_val<-median(x)
median_val
#max value
max_val<-max(x)
max_val
#min values
min_val<-min(x)
min_val
#sd values
sd_val<-sd(x)
sd_val
#variance
var_val<-var(x)
var_val
#range values
range_values<-range(x)
range_values
#mode values
table<- table(x)
table
names(table)[table == max(table)]

```

(c). Conduct the hypothesis testing Ho: 𝜇 = 60 against Ha: 𝜇 > 60 at level of significance 
𝛼 = 0:05

Answers:
Number of samples are less than 30(samples or not large) so that we can perform the t-test
Alternative hypothesis value is greater than 60 so we have to perform the 
"one tailed right sided test"

```{r}
#Answers:
#Number of samples are less than 30(samples or not large) so that we can perform the t-test
#Alternative hypothesis value is greater than 60 so we have to perform the 
#"one tailed right sided test"
mu=60
n=16
alpha=0.05
test <- t.test(x,mu=60, alternative = "greater" )
test
#from the test the p value is 0.003499 (Check the result)
test$p.value
p<-pt(test$statistic, df = n-1, lower.tail = F)
p
ifelse(p<=alpha, "The null hypothesis is rejected", "There is not enough
evidence to reject the null hypothesis.")

(or)

ifelse(test$p.value<=alpha, "The null hypothesis is rejected", "There is not enough
evidence to reject the null hypothesis.")

```
Redo part (c) at level of significance 𝛼 = 0.01. 
Compare your result with the one found in part (c)
```{r}
mu1=60
n1=16
alpha1 = 0.01
test1 <- t.test(x,mu1=60, alternative = "greater" )
test1
test1$p.value
p1<-pt(test$statistic, df = n1-1, lower.tail = F)
p1
ifelse(p1<=alpha1, "The null hypothesis is rejected", "There is not enough
evidence to reject the null hypothesis.")

(or)

ifelse(test1$p1.value<=alpha1, "The null hypothesis is rejected", "There is not enough
evidence to reject the null hypothesis.")


# in both alpha = 0.05 and 0.01, the p_value is is less than the alpha value so 
# the null hypothesis of both are rejected so alternative hypothesis is accepted

```
2. The water quality data of many dams are provided below:
(a) Make scatter plots with data for all possible pair (pairwise) (5pt)
(b) Determine Pearson’s correlation coefficient r of (a) (5pt)
(c) Interpret the results (10pt)

```{r}
#Dataset (Dams.csv) on the working directory of my system
dams <-read.csv("Dams.csv")

#pairs(dams [,c[(1:3)])
scatter.smooth(dams$Temp,dams$pH)
scatter.smooth(dams$pH, dams$TN)
scatter.smooth(dams$TN, dams$TP)

# Determine Pearson’s correlation coefficient r of (a) (5pt)
cor.test(dams$Temp,dams$pH, method = "pearson")
cor.test(dams$pH, dams$TN, method = "pearson")
cor.test(dams$TN, dams$TP, method = "pearson")

#Interpret the results (10pt)
linearmodel1<-lm(dams$Temp~dams$pH)
summary(linearmodel1)
linearmodel2<-lm(dams$pH~dams$TN)
summary(linearmodel2)
linearmodel3<-lm(dams$TN~dams$TP)
summary(linearmodel3)

summary_check1<-summary(linearmodel1)
summary_check1
linear1_coff1<-summary_check1$coefficients
linear1_coff1

summary_check2<-summary(linearmodel2)
summary_check2
linear1_coff2<-summary_check2$coefficients
linear1_coff2

summary_check3<-summary(linearmodel3)
summary_check3
linear1_coff3<-summary_check3$coefficients
linear1_coff3


```

3. Apply ANOVA for the example data (Data 1 & 2, in Exercise 7-1). 
In both applications, clearly
state a hypothesis and explain the results.
(a) Apply one-way ANOVA for Data 1. (10pt)
(b) Apply two-way ANOVA for Data 2, considering the interaction. (10pt)
```{r}
data1<-read.csv("data1.csv")
summary(data1)
boxplot(data1$pressure ~ data1$drink)
onewy_anova <- oneway.test(data1$pressure ~ data1$drink, var =T)
onewy_anova
result <-aov(data1$pressure ~ data1$drink)
result
ifelse(onewy_anova$p.value<=0.05, "Null hypothesis rejected", "Failed to reject null hypothesis")
```
(b) Apply two-way ANOVA for Data 2, considering the interaction. (10pt)
```{r}
data2<-read.csv("data2.csv")
summary(data2)
boxplot(data2$pressure ~ data2$drink*data2$gender)
result2 <-aov(data2$pressure ~ data2$drink*data2$gender)
result2
#ifelse(onewy_anova$p.value<=0.05, "Null hypothesis rejected", "Failed to reject null

```















